---
title: A Study in Factors on Musical Performance
author: Madison Loutensock
format:
  html:
    self-contained: true
    embed-resources: true
---


# Introduction
Throughout my life, I have had many experiences with private music lessons. I took private piano lessons for about 8 years and, recently, I have been working as a private piano teacher for the last 4 years. Therefore, I am very interested in analyzing what makes a lesson the most effective, so that I can be the an effective teacher.

My main research question is: "What factors in a private music lesson will help increase a student's performance?"

The sub-questions I will be focusing on are:
Do students who have a longer duration of lessons tend to get better performance scores than others?
Do students whose lessons focus on playing the instrument(practicality) tend to get better performance scores than those who focus on the theory?
Does the age of the student affect their performance score?

# Design the Study

For this analysis, I will be using data taken from a study done on Music Education Performance, where a variety of possible factors were tested, as well as an overall performance score. I will be looking at both quantitative and categorical data and I expect to do at least one t.test and an ANOVA test. The population for this analysis is all people who take private music lessons.

Response Variable: Performance Score

Possible Explanatory Variables:  
1.Duration - Quantitative  
2.Lesson Type - Categorical  
3.Age - Categorical


# Collect the Data
When I was first looking for data to analyze, my research question was "Does taking music lessons increase general education performance?" However, as I was looking into different data sets, I was drawn to this one about what affects musical performance. Due to the focus of the data being different from my original question, I had to adjust what I planned to research. I found the data through the website Kaggle, and the data was gathered in December of 2024.

# Describe/Summarize the Data

```{r}
#read in libraries and dataset
library(tidyverse)
library(mosaic)
library(rio)
library(ggplot2)
library(readr)
library(car)

setwd("C:/Users/madl1/OneDrive/Desktop/Student_Work/6-Semester_Project")

music <- read_csv("music_education_dataset.csv")
view(music)

#clean up data
unique(music$Lesson_Type)
unique(music$Age)

clean_music <- music %>%
  filter(Performance_Score >= 60,
         Performance_Score <= 100) %>%
  select(Performance_Score, Duration, Lesson_Type, Age, Stress_Level)

view(clean_music)
```

# Make Inference

### Analyzing the linear relationship between performance score and duration of play time in lessons

$$H_0:  \beta = 0$$  

$$H_a:\beta > 0$$
$$\alpha = 0.05 $$
```{r}
#explore and visualize data
ggplot(clean_music, aes(y = Performance_Score, x = Duration)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    title = "Relationship Between Duration of Play Time and Performance Scores"
  ) +
  theme_bw()
```
The scatterplot seems to have points all over the graph. There is a slight linear relationship but it is a very weak negative one.

```{r}
#correlation coefficient
cor(clean_music$Performance_Score ~ clean_music$Duration)

#testing for linear regression
lm_music <- lm(clean_music$Performance_Score ~ clean_music$Duration)
summary(lm_music)
```
__Correlation Coefficient(r)__: -0.038, which is a very weak negative relationship  
__Slope__: -0.003459  
__Intercept__: 81.679254  
__p-value__: 0.711  
__Interpretation__: The p-value is not less than alpha(0.05), therefore I fail to reject the null hypothesis. I do not have sufficient evidence to suggest that the duration of play time in lessons has an effect on a student's performance score.

```{r}
#confidence interval
confint(lm_music, level = 0.95)
```
__Confidence Interval__: I am 95% confident that the true slope of the relationship between duration and performance score is between -0.022 and 0.015. For every one second more play time in a lesson, a student's performance score changes by between -0.022 and 0.015, therefore there is little to no difference between scores in relation to the duration of play time in lessons.

```{r}
#check requirements

#residuals
qqPlot(lm_music$residuals)

#variance
plot(lm_music, which = 1)
```
__Residuals__: The residuals are normal  
__Variance__: The variance is constant  
__Trust p-value__: Yes, I can trust my p-value

### Analyzing the relationship between Lesson Type and Performance Score through independent t.test

$$H_0:  \mu_{Practical} = \mu_{Theory}$$  

$$H_a:\mu_{P} > \mu_{T}$$
$$\alpha = 0.05$$
```{r}
# explore and visualize data
favstats(clean_music$Performance_Score ~ clean_music$Lesson_Type)

ggplot(clean_music, aes(y = Performance_Score, x = Lesson_Type, fill = Lesson_Type)) +
  geom_boxplot() +
  theme_bw() +
  labs( title = "Comparing Performance Score by Lesson Type")
```
Looking at the boxplot and summary statistics, the theory focused lessons have a slightly larger median of performance scores than practical lessons. However, the variability of performance scores for practical lessons is larger, which may even it out.

```{r}
#t.test
t.test(clean_music$Performance_Score ~ clean_music$Lesson_Type, alternative = "greater")
```
__t__: 0.27839  
__df__: 93.437  
__p-value__:0.3907  
__Interpretation__: The p-value is not less than alpha, therefore I fail to reject the null hypothesis. I do not have sufficient evidence to suggest that students who take practical lessons get higher performance scores than those who take theory lessons.

```{r}
#confidence interval
t.test(clean_music$Performance_Score ~ clean_music$Lesson_Type, conf.level = 0.95)$conf.int
```
__Confidence Interval__: I am 95% confident that the true difference in performance scores of students who take practical lessons is between 4.058 less points and 5.381 more points than those who take theory lessons.

```{r}
#check for normality
qqPlot(clean_music$Performance_Score ~ clean_music$Lesson_Type)

```
__Normality__: The data is normal  
__Trust p-value__: Yes, I can trust my p-value

### Analyzing the relationship between performance score and age of the student through ANOVA

$$H_o: \mu_{10} = \mu_{11} = \mu_{12} = \mu_{13} = \mu_{14} = \mu_{15} = \mu_{16} = \mu_{17}$$
$$H_a: \text{at least one age is different from the others}$$
$$\alpha = 0.05 $$
```{r}
#explore and visualize data
favstats(clean_music$Performance_Score ~ clean_music$Age)

category_order <- c("10", "11", "12", "13", "14", "15", "16", "17")

ggplot(clean_music, aes(y = Performance_Score, x =factor(Age, levels = category_order), fill = factor(Age, levels = category_order))) +
  geom_boxplot() +
  theme_bw() +
  labs(
    x = "Age",
    title = "Comparing Performance Scores by Student Age",
    fill = "Age")
```
Looking at the summary statistics and the boxplot, age 17's median does seem to be lower than the others.

```{r}
#anova test
music_aov <- aov(clean_music$Performance_Score ~ clean_music$Age)
summary(music_aov)
```
__f__: 1.295  
__df__: 1  
__p-value__: 0.258  
__Interpretation__: The p-value is not less than alpha(0.05), therefore I fail to reject the null hypothesis. I do not have sufficient evidence to suggest that the age of the student has a significant affect on their performance score.

```{r}
#check requirements

#residuals
qqPlot(music_aov$residuals)

#standard deviation variability
sds <- favstats(clean_music$Performance_Score ~ clean_music$Age)$sd
max(sds) / min(sds)

```
__Residuals__: The residuals are normal  
__SD__: The standard deviations are approximately equal  
__Trust p-value__: Yes, I can trust my p-value

# Conclusion (Take Action)

From my findings, I conclude that neither duration of play time, lesson type, nor age have a significant affect on a student's performance score. These findings did surprise me a little, as I expected to see a relationship between at least one of them and performance score. However, my findings do imply that no matter what type of lesson a student takes, how long the lesson is, or how old they are, they can still do well in a performance. After analyzing this data, I would be interested in doing some more studies to see what factors do affect a student's performance score.
